Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-12-28 article.tex
Paper updates: finalising sections

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Resource Requirements and Robustness}
\label{sec:resource_requirements}
%%--------------------------------------------------------------------------------------------------------------------%%
This section compares resource requirements and robustness of the Jacobian-free Newton-Krylov and segregated approaches on the cases presented in Section \ref{sec:accuracy}.
Specifically, time and memory requirements are analysed, and whether the solution approach converged for all meshes and loading steps.
%In some cases, time and memory requirements from finite element software \hl{Abaqus} are given for reference.
In all cases, the Jacobian-free Newton-Krylov approach used the generalised minimal residual method (GMRES) linear solver with the direct LU preconditioner for 2-D and the Hypre \emph{BoomerAMG} multigrid preconditioner for 3-D, unless stated otherwise.
In contrast, The segregated approach used the conjugate gradient linear solver and the Hypre \emph{BoomerAMG} multigrid preconditioner for all cases.
Time and memory requirements are implementation and hardware-specific; nonetheless, it is insightful to see the relative performances of the Jacobian-free Newton-Krylov and segregated approaches on the same hardware and within the same implementation framework.
Clock times and memory usage (measured with the GNU time utility) were generated using a Mac Studio with an M2 Ultra CPU on one CPU core, where the code was built with the Clang compiler (version 16.0.0).
%\hl{Present data for one hardware first, and give comparison after.}
Examination of multi-CPU-core parallelisation is left to Section \ref{sec:parallelisation}.
% where several mesh densities are examined.
%Time and memory requirements from the classic segregated approach are given for comparison, along with the requirements from commercial finite element software Abaqus (version XXX): \hl{ask Dylan to create the Abaqus results when we are happy the paper will not change}.
%The finite element approach uses a coupled solution algorithm and direct linear solver, where a Newton-Raphson method is used for resolving nonlinearities.
%\hl{Take care when comparing times between different hardware}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Effect of the Preconditioner Choice}
\label{sec:preconditioner}
%%--------------------------------------------------------------------------------------------------------------------%%
This section examines the effect of preconditioning procedure choice on the performance of the Jacobian-free Newton Krylov approach.
%In addition, the effect of the restart parameter for the GMRES linear solver is also examined.
Three choices of preconditioning procedure are compared:
\begin{itemize}
	\item \textbf{LU} - The \emph{MUltifrontal Massively Parallel sparse direct Solver} (MUMPS) \citep{MUMPS:1, MUMPS:2} LU decomposition direct solver. A direct solver is expected to be the more robust but may suffer from excessive time and memory requirements for larger numbers of unknowns.
	\item \textbf{ILU(N)} - Incomplete LU decomposition with $N$ fill-in. ILU(N) is expected to have lower memory requirements than the LU direct solver but at the expense of robustness.
	As the system of unknowns becomes larger, the number of ILU(N) iterations is expected to increase. As the fill-in factor $N$ increases, ILU(N) approaches the robustness of a LU direct solver.
	\item \textbf{Algebraic multigrid} - The Hypre Boomerang \citep{hypre} parallelised multigrid preconditioner. Multigrid approaches have the potential to offer superior performance than other methods for larger problems, with near linear scaling of time and memory requirements.
\end{itemize}

@@

The preconditioning approaches are compared in three cases: membrane i (2-D, linear elastic), elliptic plate (3-D, linear elastic), and idealised ventricle (3-D, hyperelastic).
 Figure \ref{fig:times_memory} compares the clock times and maximum memory requirements for the three preconditioning approaches as a function of degrees of freedom.
\begin{itemize}
	\item \hl{ILU1 diverges on the ventricle for the finer two meshes: re-rnng with ILU2}
	\item \hl{LU ventricle: mesh4 is running}
	\item \hl{Comment on the results: e.g. fastest, slowest, most/least iterations, etc}
	\item \hl{Provide insights, if any: e.g. why is one faster or slower}
	\item \hl{Conclusion: which preconditioner if 'best' for which situation: probably LU is best for 'small' cases and multigrid for 'large' cases}
\end{itemize}
\begin{figure}[htbp]
   \centering
	\subfigure[Clock time]
	{
   		\includegraphics[width=0.48\textwidth]{figures/time}
	}
	\subfigure[Maximum memory]
	{
   		\includegraphics[width=0.48\textwidth]{figures/memory}  
   	}
   \caption{The clock times and maximum memory usage for three preconditioning approaches (multigrid, ILU(1), and LU) as a function of degrees of freedom}
   \label{fig:times_memory}
\end{figure}
%\begin{table}[htb]
%	\centering
%		\begin{tabular}{lllll}
%			\hline
%			Case & Number of Cells & Preconditioner & Time (in s) & Memory (in MB)  \\
%			\hline 
%			Case X & 1,234 & LU & 345 & 1,234  \\
%			Case X & 12,345 & LU & 345 & 1,234  \\
%			Case X & 123,456 & LU & 345 & 1,234  \\
%			Case X & 1,234 & ILU(0) & 345 & 1,234  \\
%			Case X & 12,345 & ILU(0) & 345 & 1,234  \\
%			Case X & 123,456 & ILU(0) & 345 & 1,234  \\
%			Case X & 1,234 & Multigrid & 345 & 1,234  \\
%			Case X & 12,345 & Multigrid & 345 & 1,234  \\
%			Case X & 123,456 & Multigrid & 345 & 1,234  \\
%			Case Y & 1,234 & LU & 345 & 1,234  \\
%			Case Z & 1,234 & LU & 345 & 1,234  \\
%			\hline
%		\end{tabular}
%	\caption{Execution times and maximum memory usage for different preconditioners on varying cases and mesh densities \hl{Try other value of N for ILU?}}
%	\label{tab:preconditioner}
%\end{table}

\hl{Other linear solver settings: mention these but no need for analyses}

In GMRES, we must choose the value of the "restarts" parameter; the default in PETSc of 30 does not always seem good enough (I have been setting it to 100). We could show its affect on one case. I think the optimal choice is linked with the choice of preconditioner (a better preconditioner means we can use a small restart value) so maybe we include this with the "preconditioner choices" section.
\hl{I suggest we do not add an analysis of this: instead we can just mention it}

This can also be considered a linear solver setting. For segregated, we always use a relative tolerance of 0.1, but initial tests for JFNK show that 0.1 is too loose and 1e-3 seems better, although maybe there is no difference with using 1e-6.
\hl{I suggest we do not add an analysis of this: instead we can just mention it}

@@

Figure \ref{fig:rhie_chow} shows a comparison of the stress \hl{(or displacement)} along the line \hl{XXX -> we need to show the effect on accuracy for different values of alpha}
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.2\textwidth]{figures/rhie_chow.pdf} 
   \includegraphics[width=0.2\textwidth]{figures/placeholder.pdf} 
   \caption{Stress predictions along the line \hl{XXX} for the \hl{YYY} case using different values of the Rhie-Chow stabilisation factor $\alpha$}
   \label{fig:rhie_chow}
\end{figure}

>>>>

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Resource Requirements and Robustness}
\label{sec:resource_requirements}
%%--------------------------------------------------------------------------------------------------------------------%%
This section compares resource requirements and robustness of the Jacobian-free Newton-Krylov and segregated approaches on the cases presented in Section \ref{sec:accuracy}.
Specifically, time and memory requirements are analysed, and whether the solution approach converged for all meshes and loading steps.
%In some cases, time and memory requirements from finite element software \hl{Abaqus} are given for reference.
In all cases, the Jacobian-free Newton-Krylov approach used the generalised minimal residual method (GMRES) linear solver with the direct LU preconditioner for 2-D and the Hypre \emph{BoomerAMG} multigrid preconditioner for 3-D, unless stated otherwise.
A restart value of 30 is used for the GMRES solver, where the \emph{loose} variant of the GMRES solver is used, which retains a small number of extra solution vectors across restarts (2 in this case).
In contrast, The segregated approach used the conjugate gradient linear solver and the Hypre \emph{BoomerAMG} multigrid preconditioner for all cases.
Time and memory requirements are implementation and hardware-specific; nonetheless, it is insightful to see the relative performances of the Jacobian-free Newton-Krylov and segregated approaches on the same hardware and within the same implementation framework.
Clock times and memory usage (measured with the GNU time utility) were generated using a Mac Studio with an M2 Ultra CPU on one CPU core, where the code was built with the Clang compiler (version 16.0.0).
%\hl{Present data for one hardware first, and give comparison after.}
Examination of multi-CPU-core parallelisation is left to Section \ref{sec:parallelisation}.
% where several mesh densities are examined.
%Time and memory requirements from the classic segregated approach are given for comparison, along with the requirements from commercial finite element software Abaqus (version XXX): \hl{ask Dylan to create the Abaqus results when we are happy the paper will not change}.
%The finite element approach uses a coupled solution algorithm and direct linear solver, where a Newton-Raphson method is used for resolving nonlinearities.
%\hl{Take care when comparing times between different hardware}

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Effect of the Preconditioner Choice}
\label{sec:preconditioner}
%%--------------------------------------------------------------------------------------------------------------------%%
This section examines the effect of preconditioning procedure choice on the performance of the Jacobian-free Newton Krylov approach.
%In addition, the effect of the restart parameter for the GMRES linear solver is also examined.
Three choices of preconditioning procedure are compared:
\begin{itemize}
	\item \textbf{LU} - The \emph{MUltifrontal Massively Parallel sparse direct Solver} (MUMPS) \citep{MUMPS:1, MUMPS:2} LU decomposition direct solver. A direct solver is expected to be the more robust but may suffer from excessive time and memory requirements for larger numbers of unknowns.
	\item \textbf{ILU(N)} - Incomplete LU decomposition with $N$ fill-in. ILU(N) is expected to have lower memory requirements than the LU direct solver but at the expense of robustness.
	As the system of unknowns becomes larger, the number of ILU(N) iterations is expected to increase. As the fill-in factor $N$ increases, ILU(N) approaches the robustness of a LU direct solver. In the current section, $N = 5$ for all cases examined.
	\item \textbf{Algebraic multigrid} - The Hypre Boomerang \citep{hypre} parallelised multigrid preconditioner. Multigrid approaches have the potential to offer superior performance than other methods for larger problems, with near linear scaling of time and memory requirements.
\end{itemize}

@@

The preconditioning approaches are compared in three cases: membrane i (2-D, linear elastic), elliptic plate (3-D, linear elastic), and idealised ventricle (3-D, hyperelastic).
 Figure \ref{fig:times_memory} compares the clock times and maximum memory requirements for the three preconditioning approaches as a function of degrees of freedom.
 Examining the clock times, the LU preconditioner is seen to be faster for all meshes in the membrane i and idealised ventricle cases, while the multigrid approach is faster for all meshes in the elliptic plate cases.
The LU approach was found to be up to $5.4 \times$ faster in the membrane i case than the multigrid approach, and $1.8 \times$ faster in the idealised ventricle case.
In contrast, the multigrid was up to $1.4 \times$ faster than the LU approach in the elliptic plate case.
%A potential explanation for why LU approach was faster in the membrane i and idealised ventricle
% for the elliptic plate case
The ILU(5) approach is seen to be slowest in all cases, in addition, the ILU(5) approach diverged on the idealised ventricle finest mesh.

Regarding memory usage, below approximately $50\,000$ degrees if freedom, the memory usage is the same for all three approaches (approximately $80$ MB), corresponding to the solver's minimum memory overhead.
For greater numbers of degrees of freedom, the LU approach uses the greatest amount of memory for both 3-D cases (idealised ventricle, elliptic plate), while the ILU(5) uses the least.
In contrast, the LU approach uses the smallest amount of memory for the 2-D case (membrane i) mesh densities examined.
The rate of memory increase is seen to be much steeper for the LU approach on the 3-D cases than the multigrid and ILU(5) approaches, with the LU results for the finest meshes not shown as they required more than the maximum available memory (64 GB).
%\begin{itemize}
%	\item \hl{Ventricle finesh mesh: ILU5 diverges and LU requires too much memory}
%	\item \hl{Elliptic plate finest mesh: LU requires too much memory}
%	\item \hl{Comment on the results: e.g. fastest, slowest, most/least iterations, etc}
%	\item \hl{Provide insights, if any: e.g. why is one faster or slower}
%	\item \hl{Conclusion: which preconditioner if 'best' for which situation: probably LU is best for 'small' cases and multigrid for 'large' cases}
%\end{itemize}
\begin{figure}[htbp]
   \centering
	\subfigure[Clock time]
	{
   		\includegraphics[width=0.48\textwidth]{figures/time}
	}
	\subfigure[Maximum memory]
	{
   		\includegraphics[width=0.48\textwidth]{figures/memory}  
   	}
   \caption{The clock times and maximum memory usage for three preconditioning approaches (multigrid, ILU(5), and LU) as a function of degrees of freedom}
   \label{fig:times_memory}
\end{figure}
%\begin{table}[htb]
%	\centering
%		\begin{tabular}{lllll}
%			\hline
%			Case & Number of Cells & Preconditioner & Time (in s) & Memory (in MB)  \\
%			\hline 
%			Case X & 1,234 & LU & 345 & 1,234  \\
%			Case X & 12,345 & LU & 345 & 1,234  \\
%			Case X & 123,456 & LU & 345 & 1,234  \\
%			Case X & 1,234 & ILU(0) & 345 & 1,234  \\
%			Case X & 12,345 & ILU(0) & 345 & 1,234  \\
%			Case X & 123,456 & ILU(0) & 345 & 1,234  \\
%			Case X & 1,234 & Multigrid & 345 & 1,234  \\
%			Case X & 12,345 & Multigrid & 345 & 1,234  \\
%			Case X & 123,456 & Multigrid & 345 & 1,234  \\
%			Case Y & 1,234 & LU & 345 & 1,234  \\
%			Case Z & 1,234 & LU & 345 & 1,234  \\
%			\hline
%		\end{tabular}
%	\caption{Execution times and maximum memory usage for different preconditioners on varying cases and mesh densities \hl{Try other value of N for ILU?}}
%	\label{tab:preconditioner}
%\end{table}

%\hl{Other linear solver settings: mention these but no need for analyses}

%In GMRES, we must choose the value of the "restarts" parameter; the default in PETSc of 30 does not always seem good enough (I have been setting it to 100). We could show its affect on one case. I think the optimal choice is linked with the choice of preconditioner (a better preconditioner means we can use a small restart value) so maybe we include this with the "preconditioner choices" section.
%\hl{I suggest we do not add an analysis of this: instead we can just mention it}
%
%This can also be considered a linear solver setting. For segregated, we always use a relative tolerance of 0.1, but initial tests for JFNK show that 0.1 is too loose and 1e-3 seems better, although maybe there is no difference with using 1e-6.
%\hl{I suggest we do not add an analysis of this: instead we can just mention it}

@@

\hl{Comment on: - crashed cases; - faster/slower/most/fewest iterations}
%Figure \ref{fig:rhie_chow} shows a comparison of the stress \hl{(or displacement)} along the line \hl{XXX -> we need to show the effect on accuracy for different values of alpha}
%\begin{figure}[htbp]
%   \centering
%%   \includegraphics[width=0.2\textwidth]{figures/rhie_chow.pdf} 
%   \includegraphics[width=0.2\textwidth]{figures/placeholder.pdf} 
%   \caption{Stress predictions along the line \hl{XXX} for the \hl{YYY} case using different values of the Rhie-Chow stabilisation factor $\alpha$}
%   \label{fig:rhie_chow}
%\end{figure}
